acquisition of TEs

### 2021-01-07 Analyze_TEs.Rmd
New stacked barchart

library(ggpirate)
library(here)
library(tidyverse)
library(tidylog)
library(wordbankr)

@@

  #Defining vocabulary-based balance score
  mutate(balance_vocab = TotalNW_NonDom/TotalNW_Dom) %>%
  #Defining proportion of TE (with reference to child's own total vocab size)
  mutate(propTE = number_of_te*2/word_vocab) %>% 
  #Defining proportion of TE (with reference to CDI)
  mutate(propTE_CDI = number_of_te/672)

@@

# Create dataframe of demographics
d_participants <- keepers_ws_TE %>%
  select(baby_id, age_days, gender, 
         lang_exp_eng, lang_exp_fre, lang_exp_other, years_education) %>%
  mutate(Female = ifelse(gender == "F", 1, 0)) %>%
  summarise(n = n(),
            age.days = mean(age_days, na.rm = TRUE),
            lang_exp_eng = mean(lang_exp_eng, na.rm = TRUE),
            lang_exp_fre = mean(lang_exp_fre, na.rm = TRUE), 
            lang_exp_other = mean(lang_exp_other, na.rm = TRUE),
            years_education = mean(years_education, na.rm = TRUE),
            number_female = sum(Female))
```

@@

@@

TEprediction_percentile <- keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TotalnumberTEs_percentile, y = number_of_te)) +
  facet_grid(. ~TEgroup) +
  labs(x = "Predicted total number of TEs (based on percentile)", y = "Observed total number of TEs",
       title = "Observed vs. Predicted total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,430) +
  ylim(-10,430)

@@

TEprediction_percentile_CDI <- keepers_ws_TE_simulation %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile, TEgroup) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te, TEgroup), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
  facet_grid(. ~TEgroup) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = 
       "Observed vs. Predicted total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,500) +
  ylim(-10,500) +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

>>>>

library(ggpirate)
library(here)
library(tidyverse)
library(tidylog)
library(wordbankr)
library("ggpubr")

@@

  #Defining vocabulary-based balance score
  #mutate(balance_vocab = TotalNW_NonDom/TotalNW_Dom) %>%
  mutate(balance_vocab = TotalNW_Dom/(TotalNW_Dom+TotalNW_NonDom)) %>%
  #Defining proportion of TE (with reference to child's own total vocab size)
  mutate(propTE = number_of_te*2/word_vocab) %>% 
  #Defining proportion of TE (with reference to CDI)
  mutate(propTE_CDI = number_of_te/672)

@@

# Create dataframe of demographics
d_participants <- keepers_ws_TE %>%
  select(baby_id, age_continuous, gender, 
         lang_exp_eng, lang_exp_fre, lang_exp_other, years_education) %>%
  mutate(Female = ifelse(gender == "F", 1, 0)) %>%
  summarise(n = n(),
            age_Month = mean(age_continuous, na.rm = TRUE),
            lang_exp_eng = mean(lang_exp_eng, na.rm = TRUE),
            lang_exp_fre = mean(lang_exp_fre, na.rm = TRUE), 
            lang_exp_other = mean(lang_exp_other, na.rm = TRUE),
            years_education = mean(years_education, na.rm = TRUE),
            number_female = sum(Female)) %>%
  mutate(PercentageFemale = number_female/n)

@@

# Correlation
```{r}
Correlation <- keepers_ws_TE %>%
  select(age_continuous, balance_vocab, word_vocab, TotalNW_Dom, TotalNW_NonDom,
         number_of_te, Dom_unique_words, NonDom_unique_words, concept_vocab) %>%
  rename(Age_month = age_continuous,
         B = balance_vocab, 
         W = word_vocab,
         D = TotalNW_Dom, 
         N = TotalNW_NonDom, 
         TE = number_of_te, 
         Du = Dom_unique_words, 
         Nu = NonDom_unique_words, 
         C = concept_vocab)

mycor <- rcorr(as.matrix(Correlation), type="pearson")
mycor$r
mycor$P

library(corrplot)
mycor <- cor(Correlation) 
corrplot(mycor)
write.csv(mycor, "ObservedTE_correlations.csv", col.names=NA)
```

###### Stacked barchat: Y=Total vocab, X= individual baby (stacked: TE, unique dom, unique nondom) ######
```{r}

##Creating dataset for stacked barchart

stacked <- keepers_ws_TE %>%
  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, balance_vocab, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te")))

##Stacked barchart (sort by total number of words)

stacked %>%
  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
  geom_col(aes(fill = Type)) +
    labs(x = "Individual infant", 
       y = "Number of words produced") + 
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c"))

##Stacked barchart (sort by total number of TEs)

stacked %>%
  ggplot(aes(x = reorder(baby_id, -TE, sum), y = NumberOfWords)) +
  geom_col(aes(fill = Type)) +
    labs(x = "Individual infant", 
       y = "Number of words produced") + 
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c"))

#stacked %>%
#  mutate(bin_balance = cut_width(balance_vocab, 0.333333333333333, boundary = 0)) %>%
#  ggplot(aes(x = reorder(Baby_ID, -NumberOfWords, sum), y = NumberOfWords)) +
#  geom_col(aes(fill = Type)) +
#  facet_grid(. ~ bin_balance, labeller=labeller(bin_balance = balance_facet)) + 
#  labs(x = "Individual infant", 
#       y = "Number of words produced") + 
#  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c"))

Bal_StackedPlot <- stacked %>%
  mutate(bin_B = cut(balance_vocab, breaks = 2, 
                     labels = c("More balanced", "Less balanced"))) %>%
  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
  geom_col(aes(fill = Type)) +
  facet_grid(. ~ bin_B) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_y_continuous(limits = c(0, 1050)) +
  labs(x = "Individual infant", 
       y = "Number of concepts produced",
       title = "Least balanced: 0 - .333") + 
  theme(plot.title = element_text(hjust = 0.5))

ggsave("Output/StackedPlot_byBalance.png", Bal_StackedPlot,
       width = 10, height = 5)

```{r}
stacked <- keepers_ws_TE %>%
  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, balance_vocab, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
  #Subset balance_vocab by 3 levels 
  mutate(BalSubset = cut(balance_vocab, breaks = 3, 
                         labels=c("HighBalance", "MidRangeBalance", "LeastBalance")))

LeastBal_plot <- stacked %>%
  filter(BalSubset == "LeastBalance") %>%
  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
  geom_col(aes(fill = Type)) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_y_continuous(limits = c(0, 1050)) +
  labs(x = "Individual infant", 
       y = "Number of concepts produced",
       title = "Least balanced") + 
  theme(plot.title = element_text(hjust = 0.5))

MidRangeBal_plot <- stacked %>%
  filter(BalSubset == "MidRangeBalance") %>%
  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
  geom_col(aes(fill = Type)) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_y_continuous(limits = c(0, 1050)) +
  labs(x = "Individual infant", 
       y = "Number of concepts produced",
       title = " ") + 
  theme(plot.title = element_text(hjust = 0.5))

HighBal_plot <- stacked %>%
  filter(BalSubset == "HighBalance") %>%
  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
  geom_col(aes(fill = Type)) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_y_continuous(limits = c(0, 1050)) +
  labs(x = "Individual infant", 
       y = "Number of concepts produced",
       title = "High balance") + 
  theme(plot.title = element_text(hjust = 0.5))

Bal_StackedPlot <- ggarrange(LeastBal_plot, 
          MidRangeBal_plot + rremove("ylab"), 
          HighBal_plot + rremove("ylab"),
          ncol = 3, nrow = 1,
          common.legend = TRUE, legend="bottom")

ggsave("Output/StackedPlot_byBalance.png", Bal_StackedPlot,
       width = 10, height = 5)

## Linear regression: TE ~ B + D + N
```{r}
TE_input_lm <- lmer(number_of_te ~ 0 + balance_vocab + TotalNW_Dom + TotalNW_NonDom,
                  data = keepers_ws_TE)

summary(TE_input_lm)
```

## Linear regression: TE ~ Du * Nu
```{r}
TE_output_lm <- lm(number_of_te ~ 0 + Dom_unique_words * NonDom_unique_words,
                  data = keepers_ws_TE)

summary(TE_output_lm)

### plot the Du*Nu interaction
#effect_plot(TE_InputConstant, pred = B, plot.points = TRUE)
interaction_DuNu <- effect(term = "Dom_unique_words:NonDom_unique_words", mod = TE_output_lm) %>%
  data.frame()

plot_interactionDuNu <- ggplot(interaction_DuNu, 
                               aes(x = Dom_unique_words, y = fit, group = NonDom_unique_words)) +
  geom_line(aes(color = NonDom_unique_words)) +
  labs(x = 'Unique dominant words',
       y = 'Number of TEs',
       color = 'Unique non-dominant words')

  ylim(-25, 600) +
  labs(x = 'Balance', y = 'Number of TEs') +
  theme_minimal()

@@

TEprediction_percentile <- keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TotalnumberTEs_percentile, y = number_of_te)) +
#  facet_grid(. ~TEgroup) +
  labs(x = "Simulated total number of TEs (based on percentile)", y = "Observed total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,430) +
  ylim(-10,430) +
  theme_minimal()

@@

TEprediction_percentile_CDI <- keepers_ws_TE_simulation %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile, TEgroup) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te, TEgroup), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
#  facet_grid(. ~TEgroup) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = 
       "Observed vs. Predicted total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,500) +
  ylim(-10,500) +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))
